The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-01 paper.tex
Proof-reading of section 3 and 4. Many little changes

\subsection{A few results from the field}

@@

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts of the non-human agent, namely the anthropomorphic form of it.
Using the term \textit{form} to describe the design of an artifact, we refer 
not only to the shape (how it looks like) but to the total 
expression of the artifact \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. The behavioral aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}.
The use of nature-inspired and in particular anthropomorphic (human-like) form in the 
design of non-human things is as old and manifold as people's tendency to anthropomorphize. 
Anthropomorphic (and life-like) forms have a long-standing tradition in design in general, and in the design of daily life things, in particular 
\citep{norman_design_1988,guthrie_bottles_2000}. Interestingly, from its very beginning robots were resembling a human being in their appearance. Starting from when the term \texit{robot} first appeared in a play by Karel \^{C}apek in the 1920ies, robots were man-made beings, created as the \textit{``shadow of man''} \citep{replicant_????}.

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance). Though conflicts were found, the HRI community more or less accepted viewing anthropomorphism as a static feature that once
observed during a short-term interaction reflects a sustaining social effect. So far, it was also measured as such. But there are some hints that this is deficient.

For instance, \cite{kanda_interactive_2004} point out that relationships among people evolve over time and similarly, also people’s attitude toward technological artifacts
and their relationship with them might evolve over time. In their study, \cite{kanda_interactive_2004} were also among the first to describe empirically \emph{novelty effects} during a long-term study with an interactive robot in a school.

There are also hints from a more theoretical side, that make us think that anthropomorphism is likely to change over 
time and with growing experience that a person has with a robot.

%effectance motivation reduces uncertainty and consequently the tendency to anthropomorphize
For instance, let's have a look at the processes that underlie anthropomorphism.
In its essence anthropomorphism means reasoning about and perceiving something
non-human (and unknown or unfamiliar) based on one's representation of the
familiar and well-known concept of being human (or one-self). According to \citep{epley_when_2008} the basic
operations that underlie these inductive inferences are the acquisition of knowledge, the
activation or elicitation of knowledge, and the application of activated
knowledge at the time of the judgment. By applying knowledge to a given situation / interaction, it is implied that one attempts to correct, adjust, or maybe
integrate less accessible information into a more automatically activated
default representation. \cite{epley_when_2008} argues that this process can be seen as a process of correction, and
thus, it is often insufficient leaving final judgments biased in the direction
of the initially activated representation. Refinements to the initial representation mean something evolves. This can be due to the fact that a person's knowledge base changes
constantly with newly acquired things, or growing experiences with a
robot, \etc. . Consequently, it is likely that the ``need to anthropomorphize'' a robot decreases
over time (from the human-centered point of view), since the user's knowledge about the robot evolves, and also the user familiarizes herself with the robot. Consequently, the robot should become more
predictable and more familiar to the human user and inferences about it can be
made based on the acquired knowledge.

In other words, the argumentation is that humans have a motivation to interact effectively in their environment \textit{(effectance)} \citep{epley_when_2008}. This means that humans are motivated to be able to understand and predict the environment and agents around them. Through this process, uncertainty is reduced, and consequently, a first unfamiliar robot, can be explained with means that are different from anthropomorphism.

%expectations
Further,

\hl{Julia is working on this section}

%\textbf{In-the-moment vs. reflective perspectives on agency}\\
%Perception of agency can be treated as anthropomorphism.
%\citep{takayama_perspectives_2012} proposes to differentiate between a \textit{in-the-moment} and \textit{reflective} perspective to better understand how human users perceive agency in robots.\footnote{``The other important distinction between the two perspectives is the \hl{cognitive processes} that they engage.''} According to Takayama, an \textit{in-the-moment} perspective refers to one's most immediate (sometimes visceral) sense in a situation. In contrast, a \textit{reflective} perspective refers to one's sense of a situation upon more distanced cogitation and consideration. For instance, we reflectively might not perceive agency in a social robot, but it can feel quite differently in the moment of interaction. A critical dimension of agency is the perspective from which something seems to have agency. \cite{takayama_perspectives_2012} argues that neglecting to separate reflective perspectives from in-the-moment perspectives of agency is one of the major sources of confusion when people talk and write about anthropomorphism. There seems to be a disconnection between what people consciously perceive and how they respond to stimuli that they may not consciously perceive.
%That means, in the initial phase of interacting with a novel device \textbf{(initialization)} people might not consciously respond but rather mindlessly \citep{nass_machines_2000}. Only after some time of \textbf{familiarization}, they might respond in a more reflective manner. This can be illustrated by the fact that people tend to deny interacting with computational systems as if they were people and yet they respond to computers in many ways that are remarkably similar to how they respond to people \citep{reeves_media_1996}.
%\cite{takayama_perspectives_2012} also applies a cognitive viewpoint on the two different perspectives of in-the-moment \vs reflective to illustrate differences. She states that in-the-moment perceptions of agency are largely shaped by bottom-up perceptual processes, evoking very immediate responses (\eg to the cues emitted by the artifact's design, see Section \ref{sec:anthropomorphic-design}).
%In contrast, according to Takayama, reflective perceptions are more often shaped by top-down processes because of the nature of reflective thought.

@@

The phenomenological model of anthropomorphism we propose
(Figure~\ref{fig:dynamics}~\citep{lemaignan2014dynamics}) represents how the
level of anthropomorphic effects~\Ant[R,H,C,t] (\ie observable manifestations of
anthropomorphism, as defined in section~\ref{sec:intro}) evolves over a
long-term human-robot interaction.

@@

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects} $\antNorm[R,H,C,t] = \frac{\ant[R,H,C,t]}{\antMax}$: because anthropomorphic effects are
not quantified on an absolute scale, we present them as a normalized value, that
spans from a minimum (no anthropomorphic effects: $\ant=0$) to a maximum (noted
\AntMax, with $t_{max} = \operatorname{arg\,max}_t \, \ant[t]$). For readability
reasons, we however omit the hat on \Ant in the following sections. We discuss
at section~\ref{measuring} possible methodologies for the quantitative and
qualitative assessment of anthropomorphic effects.

As pictured on Figure~\ref{fig:dynamics}, we suggest that $t_{max}$ actually
corresponds to the novelty effect peak. The actual maximum value of
anthropomorphic effects depends on each unique combination of human, robot and
several other factors we introduce below, and thus varies. The general
\emph{shape} of the model remains however the same and depicts the evolution of
anthropomorphism over time, \ie the general dynamics of anthropomorphism.

The model takes into account the duration of the interaction, the nature of the
interaction, as well as acquired experience and familiarization mechanisms. We
also formally introduce a so-called \emph{novelty effect} that models the first
phase of human-robot interaction, during which a specific increase of
anthropomorphic interactions is observed. We focus on \emph{long-term
interaction}, \ie direct (non-mediated), repeated interaction with the same
robot, over an extended period of time (typically longer than a week). The nature 
of the interaction (goal-directed, entertainment, etc.) may
however vary.

@@

% background shading
\path[fill=blue!20] (0,0) rectangle (0.6,5.5);
\path[fill=green!20] (0.6,0) rectangle (5.2,5.5);
\path[fill=orange!20] (5.2,0) rectangle (9.8,5.5);
\draw(0,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Initialization};
\draw(2.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Familiarization};
\draw(6.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Duration of interaction};

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
    \emph{initialization}, \emph{familiarization} and \emph{stabilization},
    preceded by a \emph{pre-interaction} phase. In the pre-interaction phase,
    users build an \emph{initial capital of anthropomorphism} (ICA, noted \ICA).
    Once the interaction starts, the level of anthropomorphism increases due to
    the \emph{novelty effect}, and then decreases to reach a \emph{stabilized
    level of anthropomorphism} (SLA, noted \SLA). \ICA and \SLA may vary
    depending on the user, the robot and the context of interaction.  During the
    interaction, unpredicted behaviors of the robot (\emph{disruptive
    behaviors}) may lead to local increases or decreases of the level of
    anthropomorphism.}

@@

@@

The \emph{initial capital of anthropomorphism} describes the initial potential
for the robot to be anthropomorphized by the human user in a given situation.
This potential depends on several factors. It has been shown, for instance, that
some \textit{people} tend to anthropomorphize more than others, that some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and that some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews~\cite{epley_when_2008}. Also the shape and design of the robot play a
role, and the context in which the interaction takes place. Our model of
anthropomorphism takes these determinants into account and initializes the level
of anthropomorphic interactions between a human and a robot to a value that we
call \emph{initial capital of anthropomorphism}, noted \ICA. \ICA describes the
first (real or imagined) contact to a robot. In this stage of pre-interaction,
people form initial expectations toward the robot and imagine how they will use
it / interact with it.

We build \ICA on three main factors that \emph{a priori} determine the
potential that a robot will be anthropomorphized:

@@

    \item \emph{Human-centered factor}: The \textbf{personality} and individual
        traits of the human user: Psychological characteristics / determinants
        that influence a person's tendency to anthropomorphize
        artifacts~\cite{epley_seeing_2007}. Other individual traits and
        demographic aspects are comprised (\eg age, gender, cultural background,
        professional background).

    \item \emph{Robot-centered factor}: The robot's \textbf{design} and how it
        appears to the human user. Characteristics of the robot's form,
        behavior, and interaction modalities (anthropomorphic
        design)~\cite{fong_survey_2003}.

    \item \emph{Situation-centered factor}: The real or imagined
        \textbf{purpose} of the robot, including the situational context in
        which it is used, as well as the task context and role in which the
        robot is used / experienced (environmental
        context)~\cite{joosse_what_2013}.

@@

\ICA comprises some hypothetical values for each of these three factors, such
that different \ICA values can be obtained, given the robot, human user, and
environmental context\fxfatal{??}.

For \textbf{personality}, we suggest to apply Epley \textit{et al.}'s
\cite{epley_seeing_2007} psychological \textit{"Three Factor Theory"} of
anthropomorphism. For instance, children generally tend to anthropomorphize
objects more than adults. Also, a person who lacks social connection is said to
be more likely to anthropomorphize. Both aspects would increase \ICA. This
means that, one and the same robot used by a different user, can lead to a
different \ICA.

For \textbf{design}, we understand that a robot that follows an anthropomorphic
design (\eg NAO) leads to a higher \ICA than a rather functional robot (\eg
Roomba). Also, a robot that is able to display facial expressions would increase
the \ICA. However, \hl{as discussed before} a classification of the "amount" of
anthropomorphic design in a robot, as for instance suggested by Fong \textit{et
al.} \cite{fong_survey_2003}, can be difficult. Nevertheless, the important
aspect is how the robot appears to the user, \eg human-like or machine-like.
Studies have shown that an anthropomorphic robot is perceived differently to a
non-anthropomorphic robot\fxfatal{...this sentence sounds strange like that.
Need references or better integration in the paragraph}.

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined context in which a robot is used and the interaction that this usage
brings along, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in Joosse
\etal~\cite{joosse_what_2013}. The authors showed for instance that when the
same robot (NAO) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different ``personalities'' to the robot. In general,
a robot which is imagined to be used in a social, entertaining or playful
context leads to a higher \ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This idea also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\cite{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the \ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased \ICA for each of the users, due to increased
human-human social interactions (the robot might be perceived to be part of the
social interaction, and in turn attributed human-like
qualities)~\cite{baxter2013do}.

@@

\fxfatal{Julia, do you have smthg to provide an initial account of what the
novelty effect is?}

@@

The woman (E.) is reading aloud parts of the "quick start manual": \emph{"When
Roomba senses dirt, the blue dirt detect light is lit and Roomba cleans more
intensely in that area."} She laughs and turns to the robot, talking directly to
it \emph{"Today you can light up darling!"}. E. carries Roomba to the entrance
door and puts it on the floor [floor type: tiles, partially covered with little
Persian carpets], she explains \emph{"It has to start from here, this is where
you enter the house!"}. After pressing the CLEAN button, the robot quickly gets
stuck with the carpet in the entrance hall, consequently it reports error 1 [
two-tone "Uh-Oh" sound followed by a narrated voice that starts off with "Error
1"]. E. is very surprised \emph{"What, it is even talking to me?"}, lifts the
robot off and puts it back for cleaning. At several occasions, while watching
how the robot moves around and cleans, E. makes compliments to the robot:
\emph{"Nice!"} or \emph{"Good job!"} or \emph{"Bravo!"}. Once, when Roomba turns
right in front of some crumbs on the kitchen floor and moves away without
cleaning them up, E. says: \emph{"There you forgot something, what is with over
there? Look, there you still have something to clean up!"}, and she points at
where she wants Roomba to go. Further, she uses directing gestures and tells the
robot: \emph{"No, go this way! Come over here!"}. The woman keeps observing the
robot and wonders about its "intentions", asking it directly: \emph{"Why do you
leave the kitchen? There is still something there on the fringe!"}. E. turns to
the experimenter, and tells her: \emph{"She [the robot] loves the hall!"}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender (\emph{"darling"}
/ \emph{"she"}); praising the robot (\emph{"Good job!"}); using pointing
gestures toward the robot, or try to direct it around with commands; attributing
a personal preference / intention to the robot (\emph{"she loves the hall!"}).

During the following visits at the woman's place, she did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. This is why we
suspect the observed effects were due to the novelty of the robot.

@@

We proposed that \ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence \SLA.
In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomorphize~\cite{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

The robot's design, on the other hand, plays a more subtle role, and strong
initial anthropomorphic design does not mandate high \SLA: lasting
anthropomorphic effects have been observed on non-anthropomorphic robots (like
the iRobot Roomba~\cite{fink_living_2013} or the military iRobot
PackBot\footnote{Rodney Brooks has reported in keynotes that occasionally
soldiers would give a name to \emph{their} PackBot and require it to be repaired
instead of being replaced by another one in case of incident.}), and on the
contrary, anthropomorphic designs can lead to higher expectation deceptions,
resulting in the robot not being used anymore.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		PART 3: EXPLANATIONS -- COGNITIVE CORRELATES, NEUROSCIENCE VIEW
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% CLAIRE & SEVERIN:
%
%

\section{Cognitive Model}
\label{sec:cognition-neuroscience}

In this section are discussing more speculative aspects and hypotheses that follow
from the model. These aspects would benefit further investigations and
experimental support.
%

%In this section we are briefly reviewing results on anthropomorphism from the cognitive neuroscience point of view.
%We then discuss more speculative aspects and hypotheses that follow
%from the model. These aspects would benefit further investigations and
%experimental support.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Cognitive model}
\label{sec:cognitive-model}

@@

\draw (-0.3,1) node[rotate=90] {Level 1}
      (-0.3,3) node[rotate=90] {Level 2}
      (-0.3,5) node[rotate=90] {Level 3};

@@

\end{tikzpicture}
%}
\caption{The three cognitive levels of anthropomorphism: Level 1 is the instinctive,
sub-cognitive identification of living peers. {\it Empathy} is characteristic
of this stage. After longer observation or short, non-contextualized interaction
(typically, a lab environment), the user reaches Level 2: the user projects a
mental model he/she is already familiar with onto the robot. After longer {\it
contextualized} interaction (typically, at home), the user enters Level 3 of
anthropomorphism: the user recomposes an accurate mental model of the robot,
based on experience. This leads to adapted interaction modalities, that may
still be anthropomorphic, or not.}
\label{fig:cognitivemodel}
\end{figure}

We propose three different cognitive phases (Figure~\ref{fig:cognitivemodel}),
which do not directly match the previously presented three phases of
anthropomorphism but are still related.

\subsubsection{Cognitive Processes and Phases}

The main underlying cognitive process in anthropomorphism is understood as
perceiving and reasoning about something non-human and unfamiliar based on one's
representation of the familiar and well-known concept of being
human~\cite{epley_when_2008}. This led us to interpret the phases of
anthropomorphic interactions as parallel cognitive phases
(Figure~\ref{fig:cognitivemodel}).

>>>>

\subsection{A few results from the field}
\label{sec:field-results}

@@

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts of the non-human agent, namely the anthropomorphic form of it.
Using the term \textit{form} to describe the design of an artifact, we refer 
not only to the shape (how it looks like) but to the total 
expression of the artifact \citep{bartneck_shaping_2004}. \cite{disalvo_hug:_2003} 
argues that \textit{form} includes a product's physical shape, materials, and 
behavioral qualities. The behavioral aspect is of special importance for interactive 
systems that show some degree of autonomy (independence).
Several categories of anthropomorphic form can be classified, but there is no 
common agreement on the borders and transitions between these categories: 
anthropomorphic, zoomorphic, caricatured, functional \citep{fong_survey_2003}.
The use of nature-inspired and in particular anthropomorphic (human-like) form in the 
design of non-human things is as old and manifold as people's tendency to anthropomorphize. 
Anthropomorphic (and life-like) forms have a long-standing tradition in design in general, and in the design of daily life things, in particular 
\citep{norman_design_1988,guthrie_bottles_2000}. Interestingly, from its very
beginning robots were resembling a human being in their appearance. Starting
from when the term \textit{robot} first appeared in a play by Karel \^{C}apek in the 1920ies, robots were man-made beings, created as the \textit{``shadow of man''} \citep{replicant_????}.

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance). Though conflicts were
found\fxwarning{SL: which conflicts are we talking about?}, the HRI
community more or less accepted viewing anthropomorphism as a static feature
that once observed during a short-term interaction reflects a sustaining social
effect.

\citet{kanda_interactive_2004} point out, however, that relationships among
people evolve over time and, in a similar manner, one can expect people's attitude toward
technological artifacts and their relationship with them to evolve as well.
In their study, \citet{kanda_interactive_2004} were also among the first to
describe empirically \emph{novelty effects} during a long-term study with an
interactive robot in a school.

From a theoretical perspective, we also expect anthropomorphism to change over
time, with growing experience with a robot. As discussed in the previous
sections, anthropomorphism means in essence \emph{reasoning about and perceiving
something non-human (and unknown or unfamiliar) based on one's
representation of the familiar and well-known concept of being human (or
one-self)}. According to \citet{epley_when_2008}, the basic operations that
underlie these inductive inferences are the \emph{acquisition of knowledge}, the
\emph{activation or elicitation of knowledge}, and the \emph{application of
activated knowledge at the time of the judgment}. By applying knowledge to a
given situation or interaction, it is implied that one attempts to correct,
adjust, or maybe integrate less accessible information into a more automatically
activated default representation. Epley argues that this process can be seen as
a process of correction, and thus, it is often insufficient leaving final
judgments biased in the direction of the initially activated representation.
Refinements to the initial representation mean something evolves. This can be
due to the fact that a person's knowledge base changes constantly with newly
acquired things, or growing experiences with a robot. Consequently, it is likely
that the ``need to anthropomorphize'' a robot decreases over time (from the
human-centered point of view), since the user's knowledge about the robot
evolves, and also the user familiarizes herself with the robot. Consequently,
the robot should become more predictable and more familiar to the human user and
inferences about it can be made based on the acquired knowledge.\fxwarning{SL:
Rephrase the last part of this paragraph}

In other words, the argumentation is that humans have a motivation to interact
effectively in their environment \textit{(effectance)} \citep{epley_when_2008}.
This means that humans are motivated to be able to understand and predict the
environment and agents around them. Through this process, uncertainty is
reduced, and consequently, a first unfamiliar robot, can be explained with means
that are different from anthropomorphism.\fxwarning{SL: Paragraph unclear to me.
Rephrase it?}

In the introduction, we reported on a few selected experimental results that
illustrate how the user tendency to anthropomorphize decreased over time.
Combined with the extensive discussion on the nature and factors of
anthropomorphism that we proposed at section~\ref{sec:anthropomorphism}, we
introduced in the following section a novel qualitative model of the evolution
of anthropomorphism over time, \ie a model of the dynamics of anthropomorphism.

@@

The phenomenological model of anthropomorphism we propose
(Figure~\ref{fig:dynamics},~\citep{lemaignan2014dynamics}) represents how the
level of anthropomorphic effects~\Ant[R,H,C,t] (\ie the observable
manifestations of anthropomorphism) evolves over a long-term human-robot
interaction.

@@

In this model, anthropomorphism is quantified by a \emph{normalized level of
anthropomorphic effects} $\antNorm[R,H,C,t] = \frac{\ant[R,H,C,t]}{\antMax}$:
because anthropomorphic effects are often qualitative and difficult to quantify
on an absolute scale, we present them as a normalized value, that spans from a
minimum (no anthropomorphic effects: $\ant=0$) to a maximum (noted \AntMax, with
$t_{max} = \operatorname{arg\,max}_t \, \ant[t]$). For the sake of readability,
we will use \Ant in the following sections to refer to the normalized level of
anthropomorphic effects. Note also that we discuss at
section~\ref{sec:measuring} possible methodologies for the quantitative and
qualitative assessment of anthropomorphic effects.

The actual maximum value of anthropomorphic effects depends on each unique
combination of human, robot and several other factors we introduce below, and
thus varies (as pictured on Figure~\ref{fig:dynamics}, we suggest besides that
$t_{max}$ actually corresponds to the novelty effect peak). The general
\emph{shape} of the model remains however the same and depicts the evolution of
anthropomorphism over time, \ie the general dynamics of anthropomorphism.

@@

% background shading
\path[fill=green!20] (0,0) rectangle (0.6,5.5);
\path[fill=blue!15] (0.6,0) rectangle (5.2,5.5);
\path[fill=orange!20] (5.2,0) rectangle (9.8,5.5);
\draw(0,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Initialization};
\draw(2.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Familiarization};
\draw(6.5,5.5) node[anchor=south west, rotate=10] {\scriptsize \sc Stabilization};
% horizontal axis
\draw[->] (0,0) -- (10,0) node[anchor=north] {$t$};
\draw(5,-0.1) node[anchor=north] {\scriptsize Duration of interaction};

@@

\caption{The dynamics of anthropomorphism. We distinguish three main phases:
    \emph{initialization}, \emph{familiarization} and \emph{stabilization},
    preceded by a \emph{pre-interaction} phase. In the pre-interaction phase,
    users build an \emph{initial capital of anthropomorphism} (ICA, noted \ICA).
    Once the interaction starts, the level of anthropomorphism increases due to
    the \emph{novelty effect}, and then decreases to reach a \emph{stabilized
    level of anthropomorphism} (SLA, noted \SLA). \ICA and \SLA may vary
    depending on the user, the robot and the context of interaction.  During the
    interaction, unpredicted behaviors of the robot (\emph{disruptive
    behaviors}) may lead to local increases or decreases of the level of
    anthropomorphic effects.}

@@

\fxwarning{SL: Mention the pre-interaction phase}

@@

The \emph{initial capital of anthropomorphism} \ICA describes the initial potential
for the robot to be anthropomorphized by the human user in a given situation.
This potential depends on several factors, introduced in
section~\ref{sec:anthropomorphism}. It has been shown, for instance, that
some \textit{people} tend to anthropomorphize more than others, that some
\textit{situations} induce anthropomorphism more than others, that
\textit{children} tend to anthropomorphize more than adults, and that some
\textit{cultures} are notorious for their anthropomorphic religions and
worldviews~\citep{epley_when_2008}. Also the shape and design of the robot play a
role, like the context in which the interaction takes place. Our model of
anthropomorphism represents these factors as a compound metric \ICA. As such,
\ICA describes the level of anthropomorphic projections just prior the first
(real or imagined) contact to a robot. In this stage of pre-interaction, people
form initial expectations toward the robot and imagine how they will use it /
interact with it.

Following the discussion in section~\ref{sec:anthropomorphism}, we build \ICA on
three main factors that determine \emph{a priori} that a robot
will be anthropomorphized:

@@

    \item ~\emph{Human-centered factor} $S$: The \textbf{personality} and individual
        traits of the human user: Psychological characteristics / determinants
        that influence a person's tendency to anthropomorphize
        artifacts~\citep{epley_seeing_2007}. Other individual traits and
        demographic aspects are comprised (\eg age, gender, cultural background,
        professional background).

    \item ~\emph{Robot-centered factor} $R$: The robot's \textbf{design} and how it
        appears to the human user. Characteristics of the robot's form,
        behavior, and interaction modalities (anthropomorphic
        design)~\citep{fong_survey_2003}.

    \item ~\emph{Context-centered factor} $C$: The real or imagined
        \textbf{purpose} of the robot, including the situational context in
        which it is used, as well as the task context and role in which the
        robot is used / experienced (environmental
        context)~\citep{joosse_what_2013}.

@@

For \textbf{personality}, we suggest to apply \citet{epley_seeing_2007}
psychological \textit{3-factor theory} of anthropomorphism. For instance,
children generally tend to anthropomorphize objects more than adults. Also, a
person who lacks social connection is said to be more likely to
anthropomorphize. Both aspects would increase \ICA. This means that, one and the
same robot used by a different user, can lead to a different \ICA.

For \textbf{design}, we understand that a robot that follows an anthropomorphic
design (\eg Nao) leads to a higher \ICA than a rather functional robot (\eg
Roomba). Also, a robot that is able to display facial expressions would increase
the \ICA. Quantify the \emph{level} of anthropomorphic design in a robot, as
suggested for instance by \citet{fong_survey_2003}, can be difficult.
Nevertheless, the important aspect is how the robot appears to the user, \eg
human-like or machine-like.

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined context in which a robot is used and the interaction that this usage
brings along, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in \citet{joosse_what_2013}. The authors showed for instance that when the
same robot (Nao) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different personalities to the robot. In general,
a robot which is imagined in a social, entertaining or playful
context leads to a higher \ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This idea also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\citep{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the \ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased \ICA for each of the users, due to increased
human-human social interactions (the robot might be perceived to be part of the
social interaction, and in turn attributed human-like
qualities~\citep{baxter2013do}).

@@

\fxfatal{SL: Julia, do you have smthg to provide an initial account of what the
novelty effect is?}

@@

The woman (E.) is reading aloud parts of the ``quick start manual'': \emph{``When
Roomba senses dirt, the blue dirt detect light is lit and Roomba cleans more
intensely in that area.''} She laughs and turns to the robot, talking directly to
it \emph{``Today you can light up darling!''}. E. carries Roomba to the entrance
door and puts it on the floor [floor type: tiles, partially covered with little
Persian carpets], she explains \emph{``It has to start from here, this is where
you enter the house!''}. After pressing the CLEAN button, the robot quickly gets
stuck with the carpet in the entrance hall, consequently it reports error 1 [
two-tone ``Uh-Oh'' sound followed by a narrated voice that starts off with ``Error
1'']. E. is very surprised \emph{``What, it is even talking to me?''}, lifts the
robot off and puts it back for cleaning. At several occasions, while watching
how the robot moves around and cleans, E. makes compliments to the robot:
\emph{``Nice!''} or \emph{``Good job!''} or \emph{``Bravo!''}. Once, when Roomba turns
right in front of some crumbs on the kitchen floor and moves away without
cleaning them up, E. says: \emph{``There you forgot something, what is with over
there? Look, there you still have something to clean up!''}, and she points at
where she wants Roomba to go. Further, she uses deictic gestures and tells the
robot: \emph{``No, go this way! Come over here!''}. The woman keeps observing the
robot and wonders about its intentions, asking it directly: \emph{``Why do you
leave the kitchen? There is still something there on the fringe!''}. E. turns to
the experimenter, and tells her: \emph{``She [the robot] loves the hall!''}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender (\emph{``darling''}
/ \emph{``she''}); praising the robot (\emph{``Good job!''}); using pointing
gestures toward the robot, or try to direct it around with commands; attributing
a personal preference / intention to the robot (\emph{``she loves the hall!''}).

During the following visits at the woman's place, she did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. The observed
effects were likely due to the novelty of the robot.

@@

We proposed that \ICA is built on three factors: user's \emph{personality},
robot's \emph{design} and interaction \emph{purpose} (or \emph{interaction
context}). The user's personality and the context of use do also influence \SLA.
In particular, it appears that the user's level of acquaintance with
technologies plays an important role in long-term tendency to
anthropomorphize~\citep{fink_living_2013} (people more familiar with technology
understand, and hence predict, better the behavior of the robot, which in turn
leads them more frequently to ultimately consider the robot as a simple tool).

The robot's design, on the other hand, plays a more subtle role, and strong
initial anthropomorphic design does not mandate high \SLA: lasting
anthropomorphic effects have been observed on non-anthropomorphic robots (like
the iRobot Roomba~\citep{fink_living_2013} or the military iRobot
PackBot\footnote{Rodney Brooks has reported in keynotes that occasionally
soldiers would give a name to \emph{their} PackBot and require it to be repaired
instead of being replaced by another one in case of incident.}), and on the
contrary, anthropomorphic designs can lead to higher expectation deceptions,
resulting in the robot not being used anymore.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		PART 3: EXPLANATIONS -- COGNITIVE CORRELATES, NEUROSCIENCE VIEW
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Cognitive Correlates of Anthropomorphism}
\label{sec:cognition-neuroscience}

This section provide an initial interpretation of anthropomorphism in
terms of \emph{cognitive correlates}: how human cognition can explain the
dynamic nature of our interactions with robots, and what are the main cognitive
stages upon which researchers can build to design better, deeper human-robot
interactions.

The next sections introduces a model of the cognitive stages involved in
human-robot interaction, that can be interpreted in parallel to the model of the
dynamics of anthropomorphism.

We also discuss two specific aspects: the cognitive impact of disruptive
behaviors (section~\ref{sec:disruptive}) on mutual modelling, and how our models
relate to the \emph{In-the-moment} versus \emph{Reflective} perspectives
introduced by~\citet{takayama_perspectives_2012}.

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{A Model of the Cognitive Stages of Human-Robot Interaction}
\label{sec:cognitive-model}

@@

\draw (-0.3,1) node[rotate=90] {Stage 1}
      (-0.3,3) node[rotate=90] {Stage 2}
      (-0.3,5) node[rotate=90] {Stage 3};

@@

\end{tikzpicture}
%}
\caption{The three cognitive stages of anthropomorphism: Stage 1 is the instinctive,
sub-cognitive identification of living peers. {\it Empathy} is characteristic
of this stage. After longer observation or short, non-contextualized interaction
(typically, a lab environment), the user reaches Stage 2: the user projects a
mental model he/she is already familiar with onto the robot. After longer {\it
contextualized} interaction (typically, at home), the user enters Stage 3 of
anthropomorphism: the user composes a custom mental model of the robot,
based on experience. This leads to adapted interaction modalities, that may
still be anthropomorphic, or not.}
\label{fig:cognitivemodel}
\end{figure}

We propose three different cognitive stages (Figure~\ref{fig:cognitivemodel}),
which are related but do not directly match the three phases of
the model of the dynamics of anthropomorphism.

\subsubsection{Cognitive Processes and Stages}

The main underlying cognitive process in anthropomorphism is understood as
perceiving and reasoning about something non-human and unfamiliar based on one's
representation of the familiar and well-known concept of being
human~\citep{epley_when_2008}. This led us to interpret the phases of
anthropomorphic interactions as parallel cognitive phases
(Figure~\ref{fig:cognitivemodel}).

The \emph{stage 1} is the instinctive, pre-cognitive identification of living
peers. Studies conducted by~\citet{rosenthal-vonderputten_experimental_2013},
who investigated the neural correlates of emotional reactions of humans towards
a robot, supports the idea that humans tend to anthropomorphize robots
intuitively in this pre-cognitive way. {\it Empathy} is characteristic of this
stage~\citep{rosenthalvonderPutten2013neural}.  It is also at this stage that
automatic activation of the human's mirror neurons system  in response to
viewing the robots action and human automatic emotional responses are
contributing at a lower level to the intuitive anthropomorphization (see Section
2.2.2 \fixme{put ref}).
